Peridynamic Beams: \\ A Non-ordinary, State Based Model

### 2014-02-17 BeamPaper.tex
actually add new diagram and plot files

\documentclass[11pt]{amsart}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{color}
\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{epstopdf}
\usepackage{etoolbox}
\usepackage{tikz}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{xfrac}
\usepackage{import}
\usepackage[section]{placeins}

@@

\title{Peridynamic Beams: \\ A Non-ordinary, State Based Model}
\author{James O'Grady \and John T. Foster}
%\date{}                                           % Activate to display a given date or no date

@@

@@

\maketitle

@@

\subinputfrom{\diagrampath}{simpleBondPair.eps_tex}

@@

\subinputfrom{\diagrampath}{bondPair.eps_tex}

@@

\theta(\vstate{Y}{}{\xi},\vstate{Y}{}{\mathbf{-\xi}}) \approx \pi-\xi \frac{\partial^2 v}{\partial x^2}+\mathcal{O}(\xi^3) =  \pi-\xi \kappa +\mathcal{O}(\xi^3); 

@@

w = \omega(\xi) \alpha \left[1+\cos(\theta(\vstate{Y}{}{\xi},\vstate{Y}{}{\mathbf{-\xi}})) \right] = \omega(\xi) \alpha\frac{\xi^2}{2}(\kappa)^2 +\mathcal{O}(\xi^4).\notag

@@

\subsection{Weighting Function and Perfect Plasticity}
\label{sec:WeightFunction}
The weighting function \(\omega(\xi)\) describes the relative contribution of each bond-pair, and can be defined according to physical or mathematical considerations. 
Consider a classical Euler-Bernoulli beam in bending with curvature \(\kappa\). 
Fibers running parallel to the neutral axis of the beam are stretched in proportion to their distance from the neutral axis, with strain \(\epsilon = y\kappa\). 
If the fibers are linearly elastic, then the axial stress at each location is \(\sigma = E\epsilon = Ey\kappa\), and the contribution to supported moment \(dM = \kappa E y^2 dA\). 
By comparing the formulations for the moments carried by the Euler beam in \cref{fig:EulerBending} and those of the bond-pair beam in \cref{fig:BPBending}, we see some definite parallels.
%
\begin{figure}[h]
\subinputfrom{\diagrampath}{EulerBending.eps_tex}
\caption{Euler beam moment contribution}
\label{fig:EulerBending}
\end{figure}
%%
\begin{figure}[h]
\subinputfrom{\diagrampath}{BondPairBending_edit.eps_tex}
\caption{Bond-pair moment contribution}
\label{fig:BPBending}
\end{figure}
%
\begin{align}
M_\text{classical}&=\int_{-\frac{t}{2}}^{\frac{t}{2}} \sigma \; y \; dA && = \int_{-\frac{t}{2}}^{\frac{t}{2}} E \kappa \; y^2 \; b(y) dy\notag \\
%
M_\text{PD}&=\int_{-\delta}^{\delta} \vstate{T}{}{\xi}\; \xi \; d\xi & = \int_{-\delta}^{\delta} \alpha \frac{\sin(\Delta\theta)}{|\xi|} \; \xi \; \omega(\xi) d\xi\: & \approx \int_{-\delta}^{\delta} \alpha \kappa |\xi| \; \omega(\xi) d\xi\notag
\end{align}
%
The term \(y\) is the distance from the beam's neutral axis and \(b(y)\) is the width of the beam at that distance from the neutral axis. 
The similarity between classical and peridynamic moment formulations suggests a possible formulation for the weighting function:
%
\begin{equation}
\omega(\xi) = |\xi| b\left(y\right) \text{at}\; y=\frac{\xi}{\delta} \frac{t}{2}\notag
\end{equation}
%
\begin{figure}[h]
\minipage[t]{0.5\textwidth}
 \centering
  \subinputfrom{\diagrampath}{WeightProfile_Uniform.eps_tex}
\caption{Weight function for a beam of rectangular cross-section}
\label{fig:WeightProfileUniform}
\endminipage\hfill
\minipage[t]{0.5\textwidth}%
  \centering
  \subinputfrom{\diagrampath}{WeightProfile_Ibeam.eps_tex}
\caption{Weight function for an I-beam}
\label{fig:WeightProfileIbeam}
\endminipage
\end{figure}
%
This weight function analogizes the relative contributions of bond pairs of different lengths to the relative contributions of fibers at different distances from the centerline. 
Examples are shown in \cref{fig:WeightProfileUniform,fig:WeightProfileIbeam}.
    While this weighting function offers no advantages over a uniform weight function in the case of the linearly elastic beam, it offers a way to model advancing plasticity.

@@

\alpha\bar{\kappa}&= \alpha(\kappa + \frac{\xi^2}{6}\kappa'') \implies \notag \\

@@

%This is consistent with Eringen's observation of nonlocal modulus as Green's function of a linear differential operator in~\cite{eringen1983differential}. 
The similarity between \cref{eq:NLbending,eq:EringenMoment} is not accidental; Eringen's gradient elasticity is the solution to the integral formulation of the nonlocal stress integral equation just as the peridynamic energy is an integral function of nonlocal displacements.
It is therefore unsurprising that, like Eringen's nonlocal elasticity, the peridynamic bending model fails to predict the stiffening associated with nanoscale cantilevers.
Instead, the advantage of peridynamic models is their natural handling of discontinuities.

\section{Numerical Simulation}
\label{sec:Numerical}
\subsection{Discretized Model}
\label{sec:Discretized}
For analysis purposes, it is often useful to represent a continuum part with a discrete version. 
Discretizing the bond-pair model is primarily matter of exchanging integrals for sums. 
%
\begin{align}
%\label{eq:discreteBeamw}
w = \alpha \left[1+\cos(\theta(\vstate{Y}{}{\xi},\vstate{Y}{}{\mathbf{-\xi}})) \right] \approx \frac{\alpha}{2}\left(\frac{v(x+\xi)-2v(x)+v(x-\xi)}{\xi}\right)^2 \notag
\end{align}
%
\begin{align}
\label{eq:discretebeam}
\alpha &= \frac{c\; \Delta x}{m} ;\; c= EI ;\; m=\sum_{i=1}^n \omega(\xi_i)\xi_i^2 \implies \nonumber \\
W&=\Delta x \sum_{i=1}^n \frac{EI}{2}\left(\frac{v(x+\xi_i)-2v(x)+v(x-\xi_i)}{\xi_i}\right)^2\notag
\end{align}
%
Discretization of the original model results in the equation of motion
\begin{equation}
\rho(\mathbf{x})\mathbf{\ddot{u}}(\mathbf{x}) = \mathbf{f}(\mathbf{x})+\sum_i \omega(\xi_i)\left\{\frac{\alpha(\mathbf{x})}{|\mathbf{p}_i |}\frac{\mathbf{p}_i}{|\mathbf{p}_i |}\times\left[\frac{\mathbf{p}_i}{|\mathbf{p}_i |}\times \frac{\mathbf{q}_i}{|\mathbf{q}_i |}\right]-\frac{\alpha(\mathbf{x}+\boldsymbol{\xi}_i)}{|\mathbf{p}_i |}\frac{(-\mathbf{p}_i)}{|\mathbf{p}_i |}\times\left[\frac{(-\mathbf{p}_i)}{|\mathbf{p}_i |}\times \frac{\mathbf{r}_i}{|\mathbf{r}_i |}\right]\right\}\notag
\end{equation}
with
\begin{align}
\mathbf{p}_i &= \boldsymbol{\xi}_i+\mathbf{u}(\mathbf{x}+\boldsymbol{\xi}_i)-\mathbf{u}(\mathbf{x});\notag\\
\mathbf{q}_i &= -\boldsymbol{\xi}_i+\mathbf{u}(\mathbf{x}-\boldsymbol{\xi}_i)-\mathbf{u}(\mathbf{x});\notag\\
\mathbf{r}_i &= \boldsymbol{\xi}_i+\mathbf{u}(\mathbf{x}+2\boldsymbol{\xi}_i)-\mathbf{u}(\mathbf{x}+\boldsymbol{\xi}_i).\notag
\end{align}
and for small displacements and rotations in a uniform beam,
\begin{equation}
\rho(x)\ddot{v}(x) = f(x)+\alpha \sum_i 2\omega(\xi_i)\frac{v(x-2\xi_i)-4v(x-\xi_i)+6v(x)-4v(x+\xi_i)+v(x+2\xi_i)}{\xi_i^2}\notag
\end{equation}
Discretizing the bond-pair model as proposed above requires that nodes be evenly spaced along the entire beam, otherwise the displacement \(v(x-\xi_i)\) is ill-defined. 
For this reason, the discretization does not allow for areas of higher and lower ``resolution''. 
%
\subsection{Numerical Method}
\label{sec:NumMethod}
Model behavior was evaluated by implementing the discretized equation of motion.
The case of a beam simply supported at both ends was chosen for simplicity in both evaluation and comparison.
Boundary conditions such as clamped supports and applied moments require careful treatment to ensure both meaningful results and ease of computation.

@@

\begin{figure}[h]
  \centering
  \scalebox{.6}{\input{\plotpath/elastic_h20_g2000.pgf}}
  \caption{The uniform-load elastic beam is accurately modeled with few nodes}
  \label{fig:elastic_g2000}
\end{figure}

As an elastic-perfectly-plastic beam exceeds the elastic limit of its material, plastic zones begin to grow on the top and bottom of the beam's cross section.
This behavior is mimicked by the plasticity of the longest bond-pairs, producing the results shown in \cref{fig:eppu_h10_g2000}.
To accurately capture this phenomenon and model beam plasticity, a finer discretization is required.
\begin{figure}[h]
  \centering
  \scalebox{.6}{\input{\plotpath/eppu_h10_g2000.pgf}}
  \caption{The elastic perfectly-plastic beam requires finer discretization}
  \label{fig:eppu_h10_g2000}
\end{figure}

@@

%
\begin{figure}[h]
  \centering
  \scalebox{.6}{\input{\plotpath/unloaded2_h10_g2000_n.pgf}}
  \caption{The need for fine discretization is even more apparent when representing residual plastic deformation}
  \label{fig:ResidualPlasticityN}
\end{figure}
%
\begin{figure}[h]
  \centering
  \scalebox{.6}{\input{\plotpath/unloaded2_n1000_g2000_h.pgf}}
  \caption{Accurately modeling residual plastic deformation also requires a small horizon}
  \label{fig:ResidualPlasticityH}
\end{figure}
%
%\begin{figure}[h]
%\minipage[t]{0.5\textwidth}
% \centering
%  \scalebox{.45}{\input{\plotpath/unloaded2_h10_g2000_n.pgf}}
%\caption{N-convergence}
%\label{fig:ResidualPlasticityN}
%\endminipage\hfill
%\minipage[t]{0.5\textwidth}%
%  \centering
%  \scalebox{.45}{\input{\plotpath/unloaded2_n1000_g2000_h.pgf}}
%\caption{H-convergence}
%\label{fig:ResidualPlasticityH}
%\endminipage
%\end{figure}

@@

\begin{figure}[h]
  \centering
  \scalebox{.65}{\input{\plotpath/brittle_h10_n200.pgf}}
  \caption{A brittle beam with prescribed center displacement}
  \label{fig:brittleBeam}
\end{figure}

>>>>

\documentclass[5p,twocolumn]{elsarticle}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{color}
\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{epstopdf}
\usepackage{etoolbox}
\usepackage{tikz}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{xfrac}
\usepackage{import}
\usepackage[section]{placeins}

@@

\journal{International Journal of Solids and Structures}

@@

\begin{frontmatter}

\title{Peridynamic Beams: \\ A Non-ordinary, State Based Model}

\author[utsa]{James O'Grady\corref{cor1}}
\ead{jogrady@gmail.com}

\author[utsa]{John Foster\corref{cor2}}
\ead{john.foster@utsa.edu}

\cortext[cor1]{Principal Corresponding author}
\cortext[cor2]{Corresponding author}

\address[utsa]{The University of Texas at San Antonio, One UTSA Circle, San Antonio, TX 78249}
%\date{}                                           % Activate to display a given date or no date

@@

\begin{keyword}
Peridynamics \sep non-ordinary models
\end{keyword}

\end{frontmatter}
%\maketitle

@@

\centering
\resizebox{\linewidth}{!}{\subinputfrom{\diagrampath}{simpleBondPair.eps_tex}}

@@

\centering
\resizebox{\linewidth}{!}{\subinputfrom{\diagrampath}{bondPair.eps_tex}}

@@

\theta(\vstate{Y}{}{\xi},\vstate{Y}{}{\mathbf{-\xi}}) &\approx \pi-\xi \frac{\partial^2 v}{\partial x^2}+\mathcal{O}(\xi^3) \notag \\
&\approx  \pi-\xi \kappa +\mathcal{O}(\xi^3); 

@@

w &= \omega(\xi) \alpha \left[1+\cos(\theta(\vstate{Y}{}{\xi},\vstate{Y}{}{\mathbf{-\xi}})) \right] \notag\\
&\approx \omega(\xi) \alpha\frac{\xi^2}{2}(\kappa)^2 +\mathcal{O}(\xi^4).\notag

@@

\subsection{Weighting Function and Perfect Plasticity}
\label{sec:WeightFunction}
The weighting function \(\omega(\xi)\) describes the relative contribution of each bond-pair, and can be defined according to physical or mathematical considerations. 
Consider a classical Euler-Bernoulli beam in bending with curvature \(\kappa\). 
Fibers running parallel to the neutral axis of the beam are stretched in proportion to their distance from the neutral axis, with strain \(\epsilon = y\kappa\). 
If the fibers are linearly elastic, then the axial stress at each location is \(\sigma = E\epsilon = Ey\kappa\), and the contribution to supported moment \(dM = \kappa E y^2 dA\). 
By comparing the formulations for the moments carried by the Euler beam in \cref{fig:EulerBending} and those of the bond-pair beam in \cref{fig:BPBending}, we see some definite parallels.
%
\begin{figure}[h]
\resizebox{\linewidth}{!}{\subinputfrom{\diagrampath}{EulerBending.eps_tex}}
\caption{Euler beam moment contribution}
\label{fig:EulerBending}
\end{figure}
%%
\begin{figure}[h]
\centering
\subinputfrom{\diagrampath}{BondPairBending_edit.eps_tex}
\caption{Bond-pair moment contribution}
\label{fig:BPBending}
\end{figure}
%
\begin{align}
M_\text{E}&=\int_{-\frac{t}{2}}^{\frac{t}{2}} \sigma \; y \; dA &= \int_{-\frac{t}{2}}^{\frac{t}{2}} E \kappa \; y^2 \; b(y) dy\notag \\
%
M_\text{PD}&=\int_{-\delta}^{\delta} \vstate{T}{}{\xi}\; \xi \; d\xi &\notag \\
&= \int_{-\delta}^{\delta} \alpha \frac{\sin(\Delta\theta)}{|\xi|} \; \xi \; \omega(\xi) d\xi\: &\approx \int_{-\delta}^{\delta} \alpha \kappa |\xi| \; \omega(\xi) d\xi\notag
\end{align}
%
The term \(y\) is the distance from the beam's neutral axis and \(b(y)\) is the width of the beam at that distance from the neutral axis. 
The similarity between classical and peridynamic moment formulations suggests a possible formulation for the weighting function:
%
\begin{equation}
\omega(\xi) = |\xi| b\left(y\right) \text{at}\; y=\frac{\xi}{\delta} \frac{t}{2}\notag
\end{equation}
%
\begin{figure}[h]
 \centering
  \subinputfrom{\diagrampath}{WeightProfile_Uniform.eps_tex}
\caption{Weight function for a beam of rectangular cross-section}
\label{fig:WeightProfileUniform}
\end{figure}

\begin{figure}
  \centering
  \subinputfrom{\diagrampath}{WeightProfile_Ibeam.eps_tex}
\caption{Weight function for an I-beam}
\label{fig:WeightProfileIbeam}
\end{figure}
%
This weight function analogizes the relative contributions of bond pairs of different lengths to the relative contributions of fibers at different distances from the centerline. 
Examples are shown in \cref{fig:WeightProfileUniform,fig:WeightProfileIbeam}.
    While this weighting function offers no advantages over a uniform weight function in the case of the linearly elastic beam, it offers a way to model advancing plasticity.

@@

\label{eq:NLbending}
\alpha\bar{\kappa}&= \alpha(\kappa + \frac{\xi^2}{6}\kappa'') \implies  \\

@@

The similarity between \cref{eq:NLbending,eq:EringenMoment} is not accidental; Eringen's gradient elasticity is the solution to the integral formulation of the nonlocal stress integral equation just as the peridynamic energy is an integral function of nonlocal displacements.
It is therefore unsurprising that, like Eringen's nonlocal elasticity, the peridynamic bending model fails to predict the stiffening associated with nanoscale cantilevers.
Instead, the advantage of peridynamic models is their natural handling of discontinuities.

\section{Numerical Simulation}
\label{sec:Numerical}
\subsection{Discretized Model}
\label{sec:Discretized}
For analysis purposes, it is often useful to represent a continuum part with a discrete version. 
Discretizing the bond-pair model is primarily matter of exchanging integrals for sums. 
%
\begin{align}
%\label{eq:discreteBeamw}
w &= \alpha \left[1+\cos(\theta(\vstate{Y}{}{\xi},\vstate{Y}{}{\mathbf{-\xi}})) \right] \notag \\
&\approx \frac{\alpha}{2}\left(\frac{v(x+\xi)-2v(x)+v(x-\xi)}{\xi}\right)^2 \notag
\end{align}
%
\begin{align}
\label{eq:discretebeam}
\alpha &= \frac{c\; \Delta x}{m} ;\; c= EI ;\; m=\sum_{i=1}^n \omega(\xi_i)\xi_i^2 \implies \nonumber \\
W&=\Delta x \sum_{i=1}^n \frac{EI}{2}\left(\frac{v(x+\xi_i)-2v(x)+v(x-\xi_i)}{\xi_i}\right)^2\notag
\end{align}
%
Discretization of the original model results in the equation of motion
\begin{align}
\rho(\mathbf{x})\mathbf{\ddot{u}}(\mathbf{x}) = \mathbf{f}(\mathbf{x})&+\sum_i \omega(\xi_i)\left\{\frac{\alpha(\mathbf{x})}{|\mathbf{p}_i |}\frac{\mathbf{p}_i}{|\mathbf{p}_i |}\times \left[ \frac{\mathbf{p}_i}{|\mathbf{p}_i |}\times \frac{\mathbf{q}_i}{|\mathbf{q}_i |}\right] \right. \notag \\
& \left. -\frac{\alpha(\mathbf{x}+\boldsymbol{\xi}_i)}{|\mathbf{p}_i |}\frac{(-\mathbf{p}_i)}{|\mathbf{p}_i |}\times\left[\frac{(-\mathbf{p}_i)}{|\mathbf{p}_i |}\times \frac{\mathbf{r}_i}{|\mathbf{r}_i |} \right] \right\} \notag
\end{align}
with
\begin{align}
\mathbf{p}_i &= \boldsymbol{\xi}_i+\mathbf{u}(\mathbf{x}+\boldsymbol{\xi}_i)-\mathbf{u}(\mathbf{x});\notag\\
\mathbf{q}_i &= -\boldsymbol{\xi}_i+\mathbf{u}(\mathbf{x}-\boldsymbol{\xi}_i)-\mathbf{u}(\mathbf{x});\notag\\
\mathbf{r}_i &= \boldsymbol{\xi}_i+\mathbf{u}(\mathbf{x}+2\boldsymbol{\xi}_i)-\mathbf{u}(\mathbf{x}+\boldsymbol{\xi}_i).\notag
\end{align}
and for small displacements and rotations in a uniform beam,
\begin{align}
&\rho(x)\ddot{v}(x) = f(x)+\alpha \sum_i 2\omega(\xi_i)\bigg( \notag\\
&\left. \frac{v(x-2\xi_i)-4v(x-\xi_i)+6v(x)-4v(x+\xi_i)+v(x+2\xi_i)}{\xi_i^2}\right) \notag
\end{align}
Discretizing the bond-pair model as proposed above requires that nodes be evenly spaced along the entire beam, otherwise the displacement \(v(x-\xi_i)\) is ill-defined. 
For this reason, the discretization does not allow for areas of higher and lower ``resolution''. 
%
\subsection{Numerical Method}
\label{sec:NumMethod}
Model behavior was evaluated by implementing the discretized equation of motion.
The case of a beam simply supported at both ends was chosen for simplicity in both evaluation and comparison.
Boundary conditions such as clamped supports and applied moments require careful treatment to ensure both meaningful results and ease of computation.

@@

\begin{figure}[h]
  \centering
  \resizebox{\linewidth}{!}{\input{\plotpath/elastic_h20_g2000.pgf}}
  \caption{The uniform-load elastic beam is accurately modeled with few nodes}
  \label{fig:elastic_g2000}
\end{figure}

As an elastic-perfectly-plastic beam exceeds the elastic limit of its material, plastic zones begin to grow on the top and bottom of the beam's cross section.
This behavior is mimicked by the plasticity of the longest bond-pairs, producing the results shown in \cref{fig:eppu_h10_g2000}.
To accurately capture this phenomenon and model beam plasticity, a finer discretization is required.
\begin{figure}[h]
  \centering
  \resizebox{\linewidth}{!}{\input{\plotpath/eppu_h10_g2000.pgf}}
  \caption{The elastic perfectly-plastic beam requires finer discretization}
  \label{fig:eppu_h10_g2000}
\end{figure}

@@

%
\begin{figure}[h]
  \centering
  \resizebox{\linewidth}{!}{\input{\plotpath/unloaded2_h10_g2000_n.pgf}}
  \caption{The need for fine discretization is even more apparent when representing residual plastic deformation}
  \label{fig:ResidualPlasticityN}
\end{figure}
%
\begin{figure}[h]
  \centering
  \resizebox{\linewidth}{!}{\input{\plotpath/unloaded2_n1000_g2000_h.pgf}}
  \caption{Accurately modeling residual plastic deformation also requires a small horizon}
  \label{fig:ResidualPlasticityH}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \resizebox{\linewidth}{!}{\input{\plotpath/brittle_h10_n200.pgf}}
  \caption{A brittle beam with prescribed center displacement}
  \label{fig:brittleBeam}
\end{figure}
